---
title: "NativeLand"
author: "Brian Yandell"
date: "2024-10-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = ".")
```

This began as a 
[Python to R Conversion](https://www.codeconvert.ai/python-to-r-converter)
from a student project.
See
[native-land.ca API](https://native-land.ca/resources/api-docs/)
for source material.
See also 
[Simple feature geometry](https://tmieno2.github.io/R-as-GIS-for-Economists/simple-feature-geometry-simple-feature-geometry-list-column-and-simple-feature.html)
from
[R as GIS for Economists](https://tmieno2.github.io/R-as-GIS-for-Economists/).

This uses
[sf](https://r-spatial.github.io/sf/)
and
[httr](https://httr.r-lib.org/)
R libraries, as well as
[tidyverse](https://www.tidyverse.org/).

This Rmarkdown file assumes that 
[nativeLandSetup.Rmd](https://github.com/byandell-envsys/landmapr/blob/main/nativeLandSetup.Rmd)
has been used to set up files.
In there, the
[native-land.ca API key](https://native-land.ca/resources/api-docs/)
is stored.
Below, calls to `readRDS("data/*.rds)` bring up saved files.

```{r}
nativeLandKey <- readRDS("data/nativeLandKey.rds")
```

Here is the complete Native Land dataset, and a US subset.
This is for internal use.

```{r}
nativeLandData <- readRDS("data/NativeLandData.rds")
nativeLandUS <- readRDS("data/nativeLandUS.rds")
nativeLandUS <- nativeLandUS[nativeLandUS$category == "territories", ]
```

### Native Land Slug (lookup names)

The `nativeLandSlug` only has `category` and `Slug`,
useful for name lookups later. 
It would be nice if `native-land.ca` had this directly.

```{r}
nativeLandSlug <- readRDS("data/NativeLandSlug.rds")
```

This file is saved locally, and used in the shiny app
[landmapr::nativeLandApp()](https://github.com/byandell-envsys/landmapr/blob/main/R/nativeLand.R).
However, it is not saved to github.
As constructed, it include only the `category` and `Slug`
columns from
[native-land.ca](https://native-land.ca/resources/api-docs/).

### Query for particlar territory, language and treaty combinations

This pulls all entries involving `akota` (Lakota and Nakota languages),
`sioux` (Oceti Sakowin Sioux) and `laramie` (Fort Laramie treaties).

```{r}
dplyr::filter(nativeLandSlug, grepl("akota|sioux|laramie", Slug))
```

### Plot Oceti Sakowin and Lakota Lands

This plots the Oceti Sakowin territories, languages and treaties
maps from `native-land.ca`, either together or on separate facets.

```{r}
str(nativeLandData)
```

```{r}
tmp <- grep("akota|oceti|laramie", nativeLandUS$Slug)
oceti_sakowin <- nativeLandUS[tmp,]
```

```{r warning = FALSE}
landmapr::ggplot_sf(color = oceti_sakowin$color) + 
  landmapr::ggplot_layer_sf(oceti_sakowin) +
  landmapr::ggplot_layer_name(oceti_sakowin) +
  ggplot2::ggtitle("Oceti Sakowin")
```

Add base map from OpenStreetMap

```{r message=FALSE}
landmapr::ggplot_sf(color = oceti_sakowin$color) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(oceti_sakowin) +
  landmapr::ggplot_layer_name(oceti_sakowin) +
  ggplot2::ggtitle("Oceti Sakowin")

```

Add facet for `category`. Note `territories` above map.

```{r message=FALSE}
landmapr::ggplot_sf(color = oceti_sakowin$color) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(oceti_sakowin) +
  landmapr::ggplot_layer_name(oceti_sakowin) +
  ggplot2::ggtitle("Oceti Sakowin") +
  ggplot2::facet_wrap(~ category)
```

### Menominee Lands

```{r}
menominee <- dplyr::filter(nativeLandData,
                           grepl("menominee", Slug))
```

```{r message=FALSE}
landmapr::ggplot_sf(color = menominee$color) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(menominee) +
  ggplot2::ggtitle("Menominee")
```

### Direct request

The `get_nativeLand` uses the `native-land.ca` API along with the key.
The challenge is you need to know the Slug exactly,
or use a lookup data frame such as `nativeLandSlug`.

```{r}
landmapr::get_nativeLand("territories", "oceti", nativeLandKey)
```

```{r message = FALSE, warning = FALSE}
landmapr::ggplot_sf(color = "blue") + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(
    landmapr::get_nativeLand("territories", "oceti-sakowin-sioux",
                             nativeLandKey)) +
  ggplot2::ggtitle("Oceti Sakowin")
```

```{r message = FALSE, warning = FALSE}
landmapr::ggplot_sf(color = "blue") + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(
    landmapr::get_nativeLand("territories", "oceti",
                             nativeLandKey, nativeLandSlug)) +
  ggplot2::ggtitle("Oceti Sakowin")
```

## Intersections with Minnesota, Wisconsin and Michigan

```{r}
census_geometry <- readRDS("data/census_geometry.rds")
```

```{r}
landmapr::ggplot_sf(color = unique(nativeLandUS$color)) + 
  landmapr::ggplot_layer_sf(dplyr::filter(nativeLandUS, category == "territories"))
```

```{r}
census_mnwimi <- dplyr::filter(
  census_geometry,
  category == "states",
  Name %in% c("Wisconsin", "Minnesota", "Michigan"))
```

```{r}
bbox_mnwimi <- sf::st_bbox(
  dplyr::filter(census_geometry, category == "states",
                Name %in% c("Wisconsin", "Minnesota", "Michigan")))
bbox_mnwimi_sf <- tibble::tibble(x = bbox_mnwimi[c("xmin","xmax")],
                     y = bbox_mnwimi[c("ymin","ymax")]) |> 
  sf::st_as_sf(coords = c("x", "y")) |> 
  sf::st_set_crs(sf::st_crs(nativeLandUS)) |>
  #get the bounding box of the two x & y coordintates, make sfc
  sf::st_bbox() |>
  sf::st_as_sfc()
```

```{r message = FALSE}
ggplot2::ggplot() +
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  ggplot2::geom_sf(data = bbox_mnwimi_sf, fill = "transparent")
```

Native lands that intersect with the (box around) US states MN, WI, MI.
Following step takes some time.

```{r}
if(file.exists(filename <- "data/nl_bbox_mnwimi.rds")) {
  nl_bbox_mnwimi <- readRDS(filename)
} else {
  nl_bbox_mnwimi <- list()
  for(i in seq_len(nrow(nativeLandData))) {
    nl_bbox_mnwimi[[i]] <- tryCatch(
      sf::st_intersection(nativeLandUS[i,], bbox_mnwimi_sf),
      error = function(e) NA)
  }
  nl_bbox_mnwimi <- dplyr::bind_rows(
    nl_bbox_mnwimi[!is.na(nl_bbox_mnwimi)])
  table(nl_bbox_mnwimi$category)
  saveRDS(nl_bbox_mnwimi, filename)
}
```

```{r warning = FALSE, message = FALSE}
landmapr::ggplot_sf(color = c(nl_bbox_mnwimi$color, census_geometry$color)) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(nl_bbox_mnwimi) +
  ggplot2::ggtitle("Native Land in MN/WI/MI") +
  landmapr:::ggplot_layer_sf(
    dplyr::filter(census_geometry, category == "states",
                  Name %in% c("Wisconsin", "Minnesota", "Michigan")))
```

Native lands that overlap with the (box around) US states MN, WI, MI.

```{r}
nl_mnwimi <- dplyr::filter(nativeLandUS, Name %in% nl_bbox_mnwimi$Name)
```

```{r warning = FALSE}
landmapr::ggplot_sf(color = c(nl_mnwimi$color, census_geometry$color)) +
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1, progress = "none") +
  landmapr::ggplot_layer_sf(nl_mnwimi) +
  ggplot2::ggtitle("Native Land in MN/WI/MI") +
  landmapr:::ggplot_layer_sf(
    dplyr::filter(census_geometry, category == "states",
                  Name %in% c("Wisconsin", "Minnesota", "Michigan"))) +
  ggplot2::facet_wrap(~ category)
```

```{r}
sf::st_drop_geometry(nl_mnwimi) |>
  dplyr::select(Slug, color, Name) |>
  dplyr::mutate(Slug = stringr::str_replace_all(Slug, "(%|[0-9])", "")) |>
  knitr::kable()
```

