---
title: "NativeLand"
author: "Brian Yandell"
date: "2024-10-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = ".")
```

This Rmarkdown file assumes that 
[nativeLandSetup.Rmd](https://github.com/byandell-envsys/landmapr/blob/main/nativeLandSetup.Rmd)
has been used to set up files.
In there, the
[native-land.ca API key](https://native-land.ca/resources/api-docs/)
is stored.
Below, calls to `readRDS(file.path(datadir, "*.rds))` bring up saved files.

The new Fort Laramie geojson files come from

- <https://sicangudata.org/repository/a/1851_treaty_bounds>
- <https://sicangudata.org/repository/a/1868_treaty_bounds>

```{r}
datadir <- "../data"
```

```{r}
nativeLandKey <- readRDS(file.path(datadir, "nativeLandKey.rds"))
```

## Fort Laramie Treaties from `native-land.ca`

```{r}
nativeLandUS <- readRDS(file.path(datadir, "nativeLandUS.rds"))
```

```{r}
laramieNL <- nativeLandUS[grepl("laramie", nativeLandUS$Slug),]
```

```{r eval=FALSE, warning = FALSE, message = FALSE}
landmapr::ggplot_sf(color = laramieNL$color) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(laramieNL) +
  landmapr::ggplot_layer_name(laramieNL) +
  ggplot2::ggtitle("native-land.ca Fort Laramie 1851 & 1868")
```

## Fort Laramie Treaties from `Ramadda`

```{r warning=FALSE}
laramieRamadda <- dplyr::bind_rows(
  "1851" = sf::st_read(file.path(datadir, "1851_treaty_bounds.geojson")) |>
    dplyr::mutate(color = "darkgreen"),
  "1868" = sf::st_read(file.path(datadir, "1868_treaty_boundaries.geojson")) |>
    dplyr::mutate(color = "darkblue"),
  .id = "Name") |>
  dplyr::select(-name) |>
  sf::st_make_valid()
laramie1851 <- dplyr::filter(laramieRamadda, Name == "1851")
laramie1868 <- dplyr::filter(laramieRamadda, Name == "1868")
```

```{r eval=FALSE, warning = FALSE, message = FALSE}
landmapr::ggplot_sf(color = c(laramieNL$color, laramieRamadda$color)) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(laramieNL) +
  landmapr::ggplot_layer_sf(laramieRamadda) +
  landmapr::ggplot_layer_name(laramieNL) +
  landmapr::ggplot_layer_name(laramieRamadda) +
  ggplot2::ggtitle("native-land.ca Fort Laramie 1851 & 1868")
```

Pull map directly from
[sicangudata.org](https://sicangudata.org)
at
<https://sicangudata.org/repository/a/1851_treaty_bounds>.

To scrape URLs from a web page using R, follow these steps:

- Install and load the necessary packages: `rvest` and `httr`.
- Use `rvest::read_html()` to read the web page content.
- Use `rvest::html_elements()` to select the anchor tags (`<a>`).
- Extract the URLs using `rvest::html_attr()` with the attribute `"href"`.
- Clean the URLs to remove any unwanted characters or fragments.
- Store the URLs in a vector or data frame for further analysis.

```{r}
sicangudata <- "https://sicangudata.org/repository/entry/get"
geojson_1851 <- "1851_treaty_bounds.geojson"
extra_1851 <- "?entryid=6e6a35ed-12e4-449b-a00b-e1f5e6dfd5c3"
url_1851 <- paste0(file.path(sicangudata, geojson_1851), extra_1851)
```

```{r}
rvest::read_html("https://www.geeksforgeeks.org//data-structures-in-r-programming") |>
  rvest::html_elements("p") |>
  rvest::html_text() |>
  print()
```

The following will not work because one cannot operate on pages in
[sicangudata.org](https://sicangudata.org), specifically on
<https://sicangudata.org/repository/a/1851_treaty_bounds>
because external activity is blocked

```{r eval=FALSE}
sic_1851 <- rvest::read_html("~/Downloads/1851 Treaty Bounds.html") |>
  rvest::html_elements("link") |>
  rvest::html_attr("href") |>
  rvest::html_text() |>
  print()
```

But one can use the specific URL (with `entryID`) to get the contents.
This requires knowing that ID, which means manually visiting the page
or having a lookup table.

```{r}
ramadda_1851 <- sf::st_read(url_1851) |>
    dplyr::mutate(color = "blue", Name = "1851") |>
  sf::st_make_valid()
```

```{r eval=FALSE, warning = FALSE, message = FALSE}
landmapr::ggplot_sf(color = c(laramieNL$color, laramieRamadda$color)) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(ramadda_1851) +
  landmapr::ggplot_layer_name(ramadda_1851) +
  ggplot2::ggtitle("Ramadda Fort Laramie 1851")
```

```{r eval=FALSE, warning = FALSE, message = FALSE}
landmapr::ggplot_sf(color = c(laramieNL$color, laramieRamadda$color)) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(laramie1851) +
  landmapr::ggplot_layer_name(laramie1851) +
  landmapr::ggplot_layer_sf(ramadda_1851) +
  landmapr::ggplot_layer_name(ramadda_1851) +
  ggplot2::ggtitle("Ramadda Fort Laramie 1851")
```

## Other Cessions from `native-land.ca`

### 1851 Laramie Treaty

```{r warning = FALSE, message = FALSE}
cessions_1851 <- landmapr:::cession_intersection(nativeLandUS, laramie1851)
landmapr::ggplot_cession(laramie1851, cessions_1851,
                         "Laramie 1851 & Cessions")
```

```{r}
landmapr::tmap_cession(laramie1851, cessions_1851)
```

```{r warning = FALSE, message = FALSE}
cessions_1851_expand <- landmapr::cession_expand(cessions_1851, nativeLandUS)
landmapr::ggplot_cession(laramie1851, cessions_1851_expand, 
                         "Expanded Laramie 1851 & Cessions")
```

```{r}
landmapr::tmap_cession(laramie1851, cessions_1851_expand)
```

### 1868 Laramie Treaty

```{r warning = FALSE, message = FALSE}
cessions_1868 <- landmapr:::cession_intersection(nativeLandUS, laramie1868)
landmapr::ggplot_cession(laramie1868, cessions_1868, "Laramie 1868 & Cessions")
```

```{r}
landmapr::tmap_cession(laramie1868, cessions_1868)
```

```{r warning = FALSE, message = FALSE}
cessions_1868_expand <- landmapr::cession_expand(cessions_1868, nativeLandUS)
landmapr::ggplot_cession(laramie1868, cessions_1868_expand, "Expanded Laramie 1868 & Cessions")
```

```{r}
landmapr::tmap_cession(laramie1868, cessions_1868_expand)
```

### Look only at Cession 701

```{r}
landmapr::ggplot_cession(
  dplyr::filter(cessions_1868_expand, Name == "701"),
  dplyr::filter(census_geometry, 
                grepl("Rosebud", Name),
                category == "aiannh"))
```

```{r}
knitr::knit_exit()
```

## Looking at SD Tribes

```{r}
census_sf <- readRDS(file.path(datadir, "census_geometry.rds"))
```

```{r}
sdTribes <- c("Rosebud", "Pine Ridge", "Crow Creek", "Cheyenne River",
              "Lower Brule", "Santee Reservation", "Standing Rock",
              "Yankton", "Flandreau")
sd_tribes <- 
  dplyr::filter(census_sf,
                grepl(paste(sdTribes, collapse = "|"), .data$Name),
                category == "aiannh")
```

```{r warning = FALSE, message = FALSE}
landmapr::ggplot_sf(color = c(sd_tribes$color, cessions$color,
                              laramieRamadda$color)) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
  landmapr::ggplot_layer_sf(sd_tribes) +
  landmapr::ggplot_layer_sf(cessions) +
  landmapr::ggplot_layer_name(cessions) +
  landmapr::ggplot_layer_name(sd_tribes) +
  ggplot2::ggtitle("native-land.ca Fort Laramie 1851 & 1868")
```

```{r warning = FALSE, message = FALSE}
landmapr::ggplot_sf(color = c(sd_tribes$color, cessions$color,
                              laramieRamadda$color)) + 
  ggspatial::annotation_map_tile(type = "osm", zoomin = -1) +
#  landmapr::ggplot_layer_sf(sd_tribes) +
  landmapr::ggplot_layer_sf(cessions) +
  landmapr::ggplot_layer_name(cessions) +
  landmapr::ggplot_layer_sf(laramieRamadda) +
  landmapr::ggplot_layer_name(laramieRamadda) +
  ggplot2::ggtitle("native-land.ca Fort Laramie 1851 & 1868")
```

```{r}
places <- landmapr:::order_places(laramieRamadda, cessions) |>
  dplyr::mutate(category = factor(Name, unique(Name)))
```

```{r}
tmap::tmap_mode("view")
landmapr::tmap_wrapper(places)
```


